### 2021-12-09 code/analysis.Rmd
new pilot now with pre-chat

---
title: "BoS pilot"
output:
  html_document: 
    toc: true
---

# TODO

# Pre-process

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

##Data import constants
data_location="data/pilot211209"

date_start=lubridate::ymd('2021-12-09')

#image_location="write-ups/images"

#model_location="code/models"
```

```{r bonus}
##This was for determining participant bonuses using the version of data with PID

# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>%
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(bonus*4/3,2)) %>%
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>% write_csv(here(data_location,"for_prolific.csv"))

```

```{r, include=F, eval=F}

d.games <- read_csv(here(data_location, 'games.csv')) %>% 
  rename(gameId = `_id`) %>% 
    filter(createdAt >= date_start)

d.players <- read_csv(here(data_location, 'players.csv')) %>% 
  select(playerId=`_id`,role=data.role)

d.chat.raw <- read_csv(here(data_location, 'rounds.csv'), guess_max=10000) %>%
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
  rename(row_id = `_id`) %>%
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>% 
  select( -ends_with('response'), -ends_with('time')) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))

d.round_results.raw <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  filter(createdAt >= date_start) %>% 
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("player", "player_", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("response", "_response", .x, fixed=T)) %>% 
  rename_with( ~ gsub("time", "_time", .x, fixed=T)) %>% 
  select(-chat) %>% 
  gather(key, value, starts_with('player')) %>% 
  separate(key, into = c('blah', 'playerId', 'info')) %>% 
  spread(info, value) %>% 
  select(-blah) %>% 
  select(-stageIds,-index,-createdAt,-submitted) %>% 
  filter(!is.na(response)) %>% 
  mutate(time=as.numeric(time)/1000) %>% 
  mutate(targets=map(targets, .f=ParseJSONColumn)) %>% 
  unnest(targets) %>% 
  mutate(image = gsub('/experiment/', '', image, fixed=TRUE),
        image = gsub('.jpeg', '', image, fixed=TRUE)) %>% 
  pivot_wider(names_from=label, names_prefix="box_", values_from=image) %>% 
    mutate(payoffs=map(payoffs, .f=ParseJSONColumn)) %>%
  unnest(payoffs) %>% 
  mutate(AA=ifelse(role=="p1", AA.p1, AA.p2),
         AB=ifelse(role=="p1", AB.p1, AB.p2),
         BA=ifelse(role=="p1", BA.p1, BA.p2),
         BB=ifelse(role=="p1", BB.p1, BB.p2)) %>% 
  write_csv(here(data_location, 'raw_results.csv'))

d.exit.survey <- read_csv(here(data_location, 'player-inputs.csv')) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c('gameId')) %>%
    rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  write_csv(here(data_location,'exit.csv'))

```

# Metrics
Paid $2.50 for 15 minutes with up to $4 bonus stated.

Sigh for those couple games.

```{r}
d.round_results <- read_csv(here(data_location,"raw_results.csv"))

summary <- d.round_results %>% group_by(repNum, gameId) %>% 
           mutate(time= time %|% 180) %>% 
  summarize(max_time=max(time)) %>% 
  group_by(gameId) %>% 
  summarize(total_time=sum(max_time)/60,
            num_rounds=max(repNum))

summary

```
```{r}

#ggplot(d.round_results, aes(x=repNum, y=time, color=playerId))+geom_point()

```

Some groups come close to maxxing bonuses

```{r}
pairs <- d.round_results %>% select(playerId,gameId) %>% unique()
d.players <- read_csv(here(data_location, 'players.csv')) %>% select(playerId=`_id`,data.bonus) %>% filter(!is.na(data.bonus)) %>% left_join(pairs) %>% arrange(gameId)

d.players
```

Number of utterances

pre-chat -- really we want word count not utterances, but eh. 
From skimming people do talk some. 
```{r}
#this is a sloppy way to split, but eh, 
chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% filter(is.na(targets))

chat %>% group_by(gameId, playerId) %>% tally()
```

```{r}
chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% filter(!is.na(targets))

chat %>% group_by(gameId, playerId) %>% tally() %>% full_join(d.players)

```

When they talk

```{r}
chat %>% group_by(gameId, playerId, repNum) %>% tally() %>% filter(n>0) %>% 
  ggplot(aes(x=repNum, y=playerId, color=gameId))+geom_point()

```
(Full chat transcript at bottom)

# What do people choose
   "AA":Best for P1
   "BB":Best for P2

Looks like there's convergence by 20 for the groups that converge, but it may not be obvious they have -- 40 rounds looks right.

```{r}
outcome <- d.round_results %>% select(gameId:repNum, response,role) %>% 
  pivot_wider(names_from=role, values_from=response) %>% 
  mutate(outcome=str_c(p1,p2))

ggplot(outcome, aes(x=repNum, y=outcome, color=gameId))+geom_jitter(height=.1)+facet_wrap(.~gameId)

```

Player 2 prefers option B, player 1 prefers action A.

```{r}
outcome <- d.round_results %>% select(gameId:repNum, response,role)

ggplot(outcome, aes(x=role, fill=response))+geom_bar(position="dodge")+facet_wrap(.~gameId)

```

# What do they say about the game

```{r}
  comments <- read_csv(here(data_location,'exit.csv')) %>% select(playerId, gameId, human, workedWell, fair, chatUseful) %>% mutate(gameId=str_sub(gameId, 0,2),playerId=str_sub(playerId, 0,2)) %>% filter(!is.na(gameId)) %>% arrange(gameId)

comments %>% select(playerId, human, workedWell)

comments %>% select(playerId,fair)

comments %>% select(playerId, chatUseful)

```
# Chat

```{r}
read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% mutate(gameId=str_sub(gameId, 0,2),playerId=str_sub(playerId, 0,2)) %>% 
  select(gameId, playerId, repNum, text) %>% kable()

```