### 2022-01-08 code/analysis.Rmd
2x2 analysis

---
title: "Study 1"
output:
  html_document: 
    toc: true
---
TODO notes:

 - get empirica time recording to behave (grumble)
 - read feedback
 - remove incomplete game(s)

# Study 1 notes

2x2 study of BoS vs PD crossed with chat or no-chat. All did 40 trials. All did a 3 minute "turing-test" free chat at the start. Run Jan 7. Target was 20 pairs in each condition.

# Pre-process

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

##Data import constants
data_location="data/study1_220107"

date_start=lubridate::ymd('2022-01-07')

#image_location="write-ups/images"

#model_location="code/models"
```

```{r bonus}
##This was for determining participant bonuses using the version of data with PID

# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>%
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(bonus*4/3,2)) %>%
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
#   filter(bonus>0) %>% write_csv(here(data_location,"for_prolific.csv"))

```

```{r}

d.treatments <- read_csv(here(data_location, 'treatments.csv')) %>% 
  rename(treatmentId=`_id`) %>% 
  select(treatmentId, name) %>% 
  separate(name, c("game_cond", "chat_cond"))

d.games <- read_csv(here(data_location, 'games.csv')) %>% 
  rename(gameId = `_id`) %>% 
  left_join(d.treatments) %>% 
    filter(createdAt >= date_start) %>% 
  select(gameId,playerIds,game_cond,chat_cond)

d.gameplayers <- d.games %>% separate(playerIds, c("p1","p2")) %>% 
  pivot_longer(p1:p2, values_to="playerId") %>% 
  select(-name)

d.players <- read_csv(here(data_location, 'players.csv')) %>% 
  select(playerId=`_id`,role=data.role, data.bonus) %>% 
  left_join(d.gameplayers)

```

```{r, include=F, eval=F}

d.chat.raw <- read_csv(here(data_location, 'rounds.csv'), guess_max=10000) %>%
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
  rename(row_id = `_id`) %>%
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>% 
  select( -ends_with('response'), -ends_with('time'), -ends_with('payoff'), -ends_with('role')) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  left_join(d.games) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))

d.round_results.raw <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  filter(createdAt >= date_start) %>% 
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("player", "player_", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("response", "_response", .x, fixed=T)) %>% 
  rename_with( ~ gsub("time", "_time", .x, fixed=T)) %>% 
  select(-chat) %>% 
  gather(key, value, starts_with('player')) %>% 
  separate(key, into = c('blah', 'playerId', 'info')) %>% 
  spread(info, value) %>% 
  select(-blah) %>% 
  select(-stageIds,-index,-createdAt,-submitted) %>% 
  filter(!is.na(response)) %>% 
  mutate(time=as.numeric(time)/1000) %>% 
  mutate(targets=map(targets, .f=ParseJSONColumn)) %>% 
  unnest(targets) %>% 
  mutate(image = gsub('/experiment/', '', image, fixed=TRUE),
        image = gsub('.jpeg', '', image, fixed=TRUE)) %>% 
  pivot_wider(names_from=label, names_prefix="box_", values_from=image) %>% 
    mutate(payoffs=map(payoffs, .f=ParseJSONColumn)) %>%
  unnest(payoffs) %>% 
  mutate(AA=ifelse(role=="p1", AA.p1, AA.p2),
         AB=ifelse(role=="p1", AB.p1, AB.p2),
         BA=ifelse(role=="p1", BA.p1, BA.p2),
         BB=ifelse(role=="p1", BB.p1, BB.p2)) %>% 
  left_join(d.games) %>% 
  write_csv(here(data_location, 'raw_results.csv'))

d.exit.survey <- read_csv(here(data_location, 'player-inputs.csv')) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c('gameId')) %>%
    rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  write_csv(here(data_location,'exit.csv'))

```

# Game length
Estimated 15 mins total including intro/exit/waiting turing etc

How long did people spend playing? (Not counting intro chat round)

```{r total-time}
# game KxJzRQeo5HikZxreo, player BkAWA2P83AnBmRcif has consistently unreasonable times (3300+, should be <180, so something's messing up)

d.round_results <- read_csv(here(data_location,"raw_results.csv"))

summary <- d.round_results %>% group_by(repNum, gameId, game_cond, chat_cond) %>% 
  filter(time<181) %>% 
  summarize(max_time=max(time)) %>% 
  group_by(gameId, game_cond, chat_cond) %>% 
  summarize(total_time=sum(max_time)/60,
            num_rounds=max(repNum)) %>% 
  filter(num_rounds==39)

ggplot(summary, aes(x=str_c(game_cond," ", chat_cond), y=total_time, color=str_c(game_cond," ", chat_cond)))+geom_jitter(width=.1)+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+
  labs(title="Game time in min", y="mins", x="condition")+theme(legend.position = "none")

summary %>% group_by(game_cond, chat_cond) %>% tally()
```
A respectable rate of game completion.

# Bonuses

Note that bonuses are not directly comparable between BoS & PD! 
But it looks like chat helps in BoS and not (much?) in PD.

```{r}
pairs <- d.round_results %>% select(playerId,gameId) %>% unique()
bonuses <- d.players %>%inner_join(pairs) %>%  select(playerId,data.bonus, chat_cond, game_cond) %>% filter(!is.na(data.bonus)) %>% left_join(pairs) %>% arrange(gameId)

ggplot(bonuses, aes(x=str_c(game_cond," ", chat_cond), y=data.bonus, color=str_c(game_cond," ", chat_cond)))+geom_jitter(width=.1)+
  labs(title="Bonus earned", y="bonus", x="condition")+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+theme(legend.position = "none")+scale_y_continuous(limits=c(0,6))
```

# Pre-chat 
```{r}
#this is a sloppy way to split, but eh, 
chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% filter(is.na(targets)) %>% 
  mutate(words=str_count(text, "\\w+")) %>%  group_by(gameId, playerId, chat_cond, game_cond) %>% summarize(words=sum(words))

ggplot(chat, aes(x=str_c(game_cond," ", chat_cond), y=words))+geom_jitter(width=.1)+
  labs(title="Words produced in pre-chat", y="words", x="condition")
```
People talked some. I haven't looked at the transcripts at all.

# During chat

This chat is raw. There are outliers that have been chopped off. 
People talk more during BoS than PD.

```{r}
game_chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(targets)) %>% filter(chat_cond=="chat") %>% 
  mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, playerId, game_cond, repNum) %>% summarize(words=sum(words))

chat_by_game <- game_chat %>% group_by(gameId, game_cond, repNum) %>% summarize(words=sum(words))

is_chat <- chat_by_game %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(game_cond, repNum) %>% summarize(pct_chat=mean(is.chat))

ggplot(game_chat, aes(x=repNum, y=words, color=game_cond))+geom_jitter(width=.1, size=1, alpha=.2)+facet_wrap(~game_cond)+scale_y_continuous(limits=c(0,30))+
  geom_smooth()+
  labs(title="Words produced per round per player", y="words")

ggplot(chat_by_game, aes(x=repNum, y=words, color=game_cond))+geom_jitter(width=.1, size=1, alpha=.2)+facet_wrap(~game_cond)+scale_y_continuous(limits=c(0,30))+
  geom_smooth()+
  labs(title="Words produced per round per game", y="words")

ggplot(is_chat, aes(x=repNum, y=pct_chat, color=game_cond))+geom_point()+
  labs(title="Pct games who said anything", y="pct")
```
Currently open questions -- is that game-related chat, or is it just that it's a friendlier game? 
Does the chatting help?

# Does chat help?

To look at if chatting helps -- we can compare forced no-chat to chat (see bonuses above). We can also compare used chat (0/1 or amount) within those that could chat.

Options for dependent variable are bonus (although there's some noise there) or option chosen.

```{r}

group_chat <- chat_by_game %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(game_cond, gameId) %>% summarize(pct_chat=mean(is.chat))

all <- d.players %>% left_join(group_chat) %>% filter(data.bonus>0)

ggplot(all, aes(x=pct_chat, y=data.bonus, color=game_cond))+geom_point()

```
Looks like chatting helps for BoS, doesn't get used much for PD.

## Choices

In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are bad for both.

In PD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. (BB is Nash equilibrium, AA is Pareto dominant.)

Not really sure how to visualize this efficiently?

Being in the chat condition v not

```{r}
outcome <- d.round_results %>% select(gameId:repNum, response,role, game_cond, chat_cond) %>% 
  pivot_wider(names_from=role, values_from=response) %>% 
  mutate(outcome=str_c(p1,p2))

ggplot(outcome, aes(x=repNum, y=outcome))+geom_jitter(width=.1, alpha=.25)+facet_grid(game_cond~chat_cond)+theme(legend.position="none")

outcome_summary <- outcome %>% group_by(game_cond, chat_cond, repNum,outcome) %>% tally() %>% group_by(game_cond, chat_cond, repNum) %>% summarize(pct=n/sum(n), outcome=outcome)

ggplot(outcome_summary, aes(x=repNum, y=pct, color=chat_cond))+geom_line()+facet_grid(game_cond~outcome)

```

Actually using the chat that round

```{r}
outcome <- d.round_results %>% select(gameId:repNum, response,role, game_cond, chat_cond) %>% 
  pivot_wider(names_from=role, values_from=response) %>% 
  mutate(outcome=str_c(p1,p2)) %>% filter(chat_cond=="chat") %>% left_join(chat_by_game) %>% mutate(is.chat=ifelse(words>0, 1,0))

outcome_summary <- outcome %>% group_by(game_cond, is.chat, repNum,outcome) %>% tally() %>% group_by(game_cond, is.chat, repNum) %>% summarize(pct=n/sum(n), outcome=outcome)

ggplot(outcome_summary, aes(x=repNum, y=pct, color=as.factor(is.chat)))+geom_line()+facet_grid(game_cond~outcome)

```