### 2022-06-14 code/text_analysis.Rmd
try add graphs

---
title: "Sbert"
output:
  html_document: 
    toc: true
---

# Pre-process

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)
library(tidytext)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

##Data import constants
data_location="data/study2_220413"

date_start=lubridate::ymd('2022-04-13')

study1_loc="data/study1_220107"
image_location="write-ups/images"

#model_location="code/models"
```

```{r comparison data}
study1.chat <- read_csv(here(study1_loc,"raw_chat.csv")) %>% 
    mutate(gametype=game_cond,
         mix="pure") %>% 
  select(gameId,text, type, repNum, playerId, chat_cond, gametype,mix, game_cond, targets)

study1.rounds <- read_csv(here(study1_loc,"raw_results.csv")) %>% 
  mutate(gametype=game_cond,
         mix="pure") %>% 
  select(gameId,repNum, playerId,payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype,mix, role)

study2.chat <- read_csv(here(data_location,"raw_chat.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,text, type, repNum, playerId, chat_cond, gametype,mix, game_cond, targets)

study2.rounds <- read_csv(here(data_location,"raw_results.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,repNum, playerId,payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype,mix, role)

rounds <- study1.rounds %>% union(study2.rounds)
chat <- study1.chat %>% union(study2.chat)

```

```{r prep}

raw_text <- chat %>% left_join(rounds) %>% 
  filter(!is.na(repNum)) %>% 
  filter(!is.na(text)) %>% 
  select(gameId,repNum,text, playerId, gametype,mix,game_cond,box_A,box_B,payoff,response,AA,AB,BA,BB) %>% 
  unnest_tokens(sentence, input=text, token="sentences") %>% 
  filter(!is.na(sentence))

#sample <- raw_text %>% filter(gameId=="3WzEi9zkHF77vdTw5") %>% write_csv("sample.csv")
pre_sbert <- raw_text %>% write_csv("pre_sbert.csv")
```

# Interlude
This is where they need to get embedding with jupyter

# Post-wrangle

```{r, include=F, eval=F}
library(reticulate)
np <- import("numpy")
mat = np$load('post_sbert.npy')
saveRDS(mat,'post_sbert.RData')
```

```{r, include=F,}
ssb <- pre_sbert %>% bind_cols(readRDS('post_sbert.RData')  %>% as_tibble())

grouped <- ssb %>%
  group_by(gameId,repNum, playerId, gametype,mix,game_cond,box_A,box_B,payoff,response,AA,AB,BA,BB) %>% 
  summarize(across(starts_with('V'), mean)) %>% 
  ungroup()

F_mat <- grouped %>% select(starts_with("V")) %>% as.matrix() #Features
M_mat <- grouped %>% select(-starts_with("V")) %>% mutate(feature_ind=row_number())
```

We average across vectors for what each person said each rep.

```{r helpers}
# note: cor expects features to be in columns so we transpose
get_sim_matrix = function(df, F_mat, method = 'cosine') {
  feats = F_mat[df$feature_ind,]
  if(method == 'cor') {
    return(cor(t(feats), method = 'pearson'))
  } else if (method == 'euclidean') {
    return(as.matrix(dist(feats, method = 'euclidean')))
  } else if (method == 'cosine') {
    return(as.matrix(lsa::cosine(t(feats))))
  } else {
    stop(paste0('unknown method', method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids) {
  ut <- upper.tri(cormat)
  data.frame(
    dim1 = ids[row(cormat)[ut]],
    dim2 = ids[col(cormat)[ut]],
    sim  = as.numeric(cormat[ut])
  ) %>%
    mutate(dim1 = as.character(dim1),
           dim2 = as.character(dim2))
}

make_within_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          .$repNum)) %>%
    mutate(rep1 = as.numeric(dim1), 
           rep2 = as.numeric(dim2)) 
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          as.character(.$combinedId)))
}
```

# Analysis

## Convergence within groups

look within a game, across all reps and playerIDs

```{r }

game_convergence <- M_mat %>% 
  group_by(gameId,game_cond) %>% 
  mutate(combinedId=str_c(playerId,repNum,sep="_")) %>% 
  make_across_df(F_mat, 'cosine') %>% 
  separate(dim1, into=c("playerId_1","repNum_1"), convert=T, sep="_") %>% 
  separate(dim2, into=c("playerId_2","repNum_2"), convert=T, sep="_") %>% 
  mutate(repDist=abs(repNum_1-repNum_2),
         earlierRep=ifelse(repNum_1>repNum_2,repNum_2,repNum_1),
         samePlayer=ifelse(playerId_1==playerId_2,1,0)) %>% 
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  filter(!is.na(samePlayer)) %>% 
  ungroup()

```

0 is not same speaker, 1 is same speaker.

```{r}
game_convergence %>% filter(repDist<5) %>% ggplot(aes(x=earlierRep, y=sim, color=repDist))+geom_point(alpha=.1)+geom_smooth()+facet_grid(game_cond~samePlayer)+labs(title="Rep distance <5")

game_convergence %>% ggplot(aes(x=repDist, y=sim, color=earlierRep))+geom_point(alpha=.01)+geom_smooth()+facet_grid(game_cond~samePlayer)

```

We're not really seeing much, similarity is higher in rounds that are closer together, but if we only look at relatively close reps (within 5 of each other), there's not much by way of pattern.

```{r, eval=F}
m_convergence <- lmer(sim ~ repDist*game_cond+samePlayer+earlierRep*game_cond + (1|gameId), data=game_convergence)

summary(m_convergence)
```

Unclear what interactions we should be including.

repDistance: reps that are further apart from each other are less similar (note this is confounded w/ earlier rep)

being said by the same person increases similarity

## Divergence between groups

```{r }

game_divergence <- M_mat %>% 
  group_by(game_cond) %>% 
  mutate(combinedId=str_c(gameId,repNum,sep="_")) %>% 
  make_across_df(F_mat, 'cosine') %>% 
  separate(dim1, into=c("gameId_1","repNum_1"), convert=T, sep="_") %>% 
  separate(dim2, into=c("gameId_2","repNum_2"), convert=T, sep="_") %>% 
  filter(gameId_1!=gameId_2) %>% 
  mutate(repDist=abs(repNum_1-repNum_2),
         earlierRep=ifelse(repNum_1>repNum_2,repNum_2,repNum_1)) %>% 
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  ungroup()
```

```{r}
game_divergence %>% filter(repDist==0) %>% ggplot(aes(x=earlierRep, y=sim))+geom_point(alpha=.1)+geom_smooth()+facet_grid(game_cond~.)+labs(title="Rep distance == 0")

game_divergence %>% ggplot(aes(x=repDist, y=sim, color=earlierRep))+geom_point(alpha=.01)+geom_smooth()+facet_grid(game_cond~.)

```

Not seeing anything useful here.

```{r, eval=F}

m_divergence <- lm(sim ~ repDist*game_cond+earlierRep*game_cond, data=game_convergence)

summary(m_divergence)
```

There's more difference the further apart the reps are, and maybe slight differences between the game types, but we're not seeing the clear divergence patterns that arise in say tangrams. (Which makes sense if groups are mostly converging on a limited number of patterns we wouldn't expect difference between them.)