### 2022-06-27 code/exp2_mixed.Rmd
data and prelim analysis of new study

---
title: "Mix of PD and BoS"
output:
  html_document: 
    toc: true
---

# Pre-process

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

##Data import constants
data_location="data/study3_220627"

date_start=lubridate::ymd('2022-06-24')

image_location="write-ups/images"

#model_location="code/models"
```

```{r bonus}
##This is for determining participant bonuses using the version of data with PID

# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>%
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(bonus*4/3,2)+) %>%
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
#   filter(bonus>0) %>% write_csv(here(data_location,"for_prolific.csv"))
# 
# cost <- read_csv(here(data_location, "player_payments.csv"))
# 
# sum(cost$cost)
```

```{r, include=F, eval=F}

d.treatments <- read_csv(here(data_location, 'treatments.csv')) %>% 
  rename(treatmentId=`_id`) %>% 
  select(treatmentId, name) %>% 
  mutate(name=ifelse(name=="mix-chat", "mix-nochat", name)) %>% #VB 220413: turns out naming conventions matter...
  separate(name, c("game_cond", "chat_cond"))

d.games <- read_csv(here(data_location, 'games.csv')) %>% 
  rename(gameId = `_id`) %>% 
  left_join(d.treatments) %>% 
    filter(createdAt >= date_start) %>% 
  select(gameId,playerIds,game_cond,chat_cond)

d.gameplayers <- d.games %>% separate(playerIds, c("p1","p2")) %>% 
  pivot_longer(p1:p2, values_to="playerId") %>% 
  select(-name)

d.players <- read_csv(here(data_location, 'players.csv')) %>% 
  select(playerId=`_id`,role=data.role, data.bonus) %>% 
  left_join(d.gameplayers)

```

```{r, include=F, eval=F}

d.chat.raw <- read_csv(here(data_location, 'rounds.csv'), guess_max=10000) %>% 
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
  rename(row_id = `_id`) %>% 
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>% 
  select( -ends_with('response'), -ends_with('time'), -ends_with('payoff'), -ends_with('role')) %>% 
  rename(data.gametype=data.type) %>% #VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  left_join(d.games) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))

d.round_results.raw <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  filter(createdAt >= date_start) %>% 
    rename(data.gametype=data.type) %>% #VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("player", "player_", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("response", "_response", .x, fixed=T)) %>% 
  rename_with( ~ gsub("time", "_time", .x, fixed=T)) %>% 
  select(-chat) %>% 
  gather(key, value, starts_with('player')) %>% 
  separate(key, into = c('blah', 'playerId', 'info')) %>% 
  spread(info, value) %>% 
  select(-blah) %>% 
  select(-stageIds,-index,-createdAt,-submitted) %>% 
  filter(!is.na(response)) %>% 
  mutate(time=as.numeric(time)/1000) %>% 
  mutate(targets=map(targets, .f=ParseJSONColumn)) %>% 
  unnest(targets) %>% 
  mutate(image = gsub('/experiment/', '', image, fixed=TRUE),
        image = gsub('.jpeg', '', image, fixed=TRUE)) %>% 
  pivot_wider(names_from=label, names_prefix="box_", values_from=image) %>% 
    mutate(payoffs=map(payoffs, .f=ParseJSONColumn)) %>%
  unnest(payoffs) %>% 
  mutate(AA=ifelse(role=="p1", AA.p1, AA.p2),
         AB=ifelse(role=="p1", AB.p1, AB.p2),
         BA=ifelse(role=="p1", BA.p1, BA.p2),
         BB=ifelse(role=="p1", BB.p1, BB.p2),
         type=ifelse(AB==0 & BA==0, "BoS", "PD")) %>% 
  left_join(d.games)

d.exit.survey <- read_csv(here(data_location, 'player-inputs.csv')) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c('gameId')) %>%
    rename_with(~ gsub("data.", "", .x, fixed = TRUE))

d.complete <- d.round_results.raw %>% group_by(gameId, playerId) %>% 
  summarize(max_rep=max(repNum)) %>% filter(max_rep==39) %>% select(gameId,playerId) %>% 
  write_csv(here(data_location,"complete.csv"))

d.chat <- d.chat.raw %>% inner_join(d.complete) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))
d.round_results <- d.round_results.raw %>% inner_join(d.complete) %>% 
  write_csv(here(data_location, 'raw_results.csv'))
d.exit <- d.exit.survey %>% inner_join(d.complete) %>% 
  write_csv(here(data_location,'exit.csv'))

```

# Read data

```{r}
chat <- read_csv(here(data_location,"raw_chat.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,text, type, repNum, playerId, chat_cond, gametype,mix, game_cond, targets)
rounds <- read_csv(here(data_location,"raw_results.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,repNum, playerId,payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype,mix, role)

```

# Summary of experiment
In the expt reported here pairs of participants  played 40 rounds of a game-theory type game. At the start, each pair had 3 minutes of free chat, and then played the game. We recruited for 20 games in chat and 20 games in no-chat conditions.

 - 4 "spiked" BoS trials where one of the rewards is high (25-30) and the other is normal (3-7) These are hard coded to appear on certain trials (7,16,23,32 / 40 trials total).
  - ~16 normal BoS trials where both rewards are from [1-9]
  - ~10 normal PD trials where cooperating is welfare maximizing ( 2 * coop payoff greater than or equal to defect payoff). All rewards are chosen from [1-12] respecting this.
  - ~10 “sacrifice” PD trials where having one person defect and the other take the sucker payoff is welfare maximizing (the defect payoff is > 2 * coop payoff). All rewards are chosen from [1-12] respecting this. 
Aside from the 4 “spiked” trials that occur at set places, the others are determined by randomly drawing from a 16:10:10 ratio distribution, so games will vary in exact proportions and locations of these trials.

```{r}

chat %>% select(gameId, game_cond, chat_cond) %>% unique() %>% group_by(game_cond, chat_cond) %>% tally()

```

Fewer chat games than non-chat games -- not sure why the differential loss -- maybe due to order of running??

## Pre-reg'd analysis plan

This is exploratory research, here are comparisons we plan to make (i.e. with graphs).

- [x] Do games with the option to chat get better scores than games without (total, and subdivided by trial type)?

- [x] Which quadrant (combination of A,B responses) do players land in as a function of chat/nochat and trial condition? (Do games with chat choose the off-diagonal option in “sacrifice” PD trials?)

- [x] How much language do players use as a function of trial number and type?

- (More qualitative) What heuristics/strategies do players use? (gleaned from post-experiment question and chat logs) How successful are they and which ones require how much language to coordinate?

- Are there signs of conventionalizing on ways of describing trials / strategies beyond names for boxes?

## other stuff to explore

- need to do intent to treat analysis
- look at what players write as their strategy
- can we tell what heuristics are being used (by trial, as function of alternating, as function of score so far?)
- [it's a mess] correlation w/i game between outcomes on different trial types

# Overall outcomes

A coarse first pass is to look at how much bonus people earned, as bonus is propotional to points.

By trial type:

```{r}
bonuses <- rounds %>%group_by(playerId,gameId, chat_cond, gametype) %>% summarize(payoff=mean(payoff))

ggplot(bonuses, aes(x=str_c(chat_cond), y=payoff, color=str_c(chat_cond)))+geom_jitter(width=.1, alpha=.5)+
  facet_grid(.~gametype)+
  scale_color_viridis(discrete=T)+
  labs(y="mean_reward", x="condition")+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.7,color="black")+theme(legend.position = "none")+
  coord_cartesian(ylim=c(0,20))+
  labs(y="Mean reward / trial")+
    theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=14),
        strip.text=element_text(size=14) )

```

Aggregate:

```{r}
rounds %>%group_by(playerId,gameId, chat_cond) %>% summarize(payoff=sum(payoff)) %>% ggplot(aes(x=chat_cond, y=payoff, color=str_c(chat_cond)))+geom_jitter(width=.1, alpha=.5)+
  scale_color_viridis(discrete=T)+
 stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.7,color="black")+theme(legend.position = "none")+
  labs(y="Total reward")+
    theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=14),
        strip.text=element_text(size=14) )

```

Clear aggregate difference based on condition.

## Pre-chat

We hope that players use the pre-chat to communicate with each other and convince themselves that the other person is a human. All games are the same at this point, so we don't expect differential amounts of talk.

```{r}
prechat <- chat %>% filter(!is.na(text)) %>% 
  filter(is.na(targets)) %>% mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, playerId, game_cond, chat_cond) %>% summarize(words=sum(words))

ggplot(prechat, aes(x=str_c(game_cond,"\n",chat_cond), y=words))+geom_jitter(width=.1, color="lightblue")+ stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+theme(legend.position = "none")+labs(y="Pre-chat words / player", x="")+
      theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=14),
        strip.text=element_text(size=14) )

```

## During chat

Chatting seems to increase slightly across the games, but many groups even in the chat condition don't use the chat.

```{r}
game_chat <- chat %>% filter(chat_cond=="chat") %>% filter(!is.na(targets)) %>% select(-type) %>% left_join(rounds) %>% 
  mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, playerId, gametype,  mix, repNum) %>% summarize(words=sum(words))

chat_by_pair <- game_chat %>% group_by(gameId, gametype, mix, repNum) %>% summarize(words=sum(words))

is_chat <- chat_by_pair %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(repNum) %>% summarize(pct_chat=mean(is.chat))
ggplot(is_chat, aes(x=repNum, y=pct_chat))+geom_point()+geom_smooth()+
  labs(y="pct of games who said anything each round")
```

# Are they human?

```{r bot}

exit <- read_csv(here(data_location,'exit.csv'))

exit %>% 
  group_by(game_cond, chat_cond, human) %>% 
  tally() %>% 
  pivot_wider(names_from=human, values_from=n) %>% 
  mutate(pct=yes/(no+yes))

```

Yay, they mostly think they're human!!

# Chat and payoff

We look at how many rounds had non-zero chat (of those with the option to chat) compared to mean payoff.

```{r}
overall_success <- rounds %>% filter(chat_cond=="chat") %>% select(playerId, gameId, gametype, payoff, repNum) %>% inner_join(chat_by_pair) %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(gameId) %>% summarize(mean_payout=mean(payoff), mean_chat=mean(is.chat), mean_words=mean(words))

ggplot(overall_success, aes(x=mean_chat, y=mean_payout))+geom_point()+geom_smooth(method="lm")+labs(y="Mean Payout", x="Proportion round with nonzero chat")

ggplot(overall_success, aes(x=mean_words, y=mean_payout))+geom_point()+geom_smooth(method="lm")+labs(y="Mean payout", x="Average words / round")

```

Using the chat more of the time, and using more words are both associated with higher payoffs.

We can break this down by trial type.

```{r}
success <- rounds %>% filter(chat_cond=="chat") %>% select(playerId, gameId, gametype, payoff, repNum) %>% inner_join(chat_by_pair) %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(gameId, gametype) %>% summarize(mean_payout=mean(payoff), mean_chat=mean(is.chat), mean_words=mean(words))

ggplot(success, aes(x=mean_chat, y=mean_payout, color=gametype))+geom_point()+geom_smooth(method="lm")+labs(y="Mean Payout", x="Proportion round with nonzero chat")

ggplot(success, aes(x=mean_words, y=mean_payout, color=gametype))+geom_point()+geom_smooth(method="lm")+labs(y="Mean payout", x="Average words / round")

```

Chat helps a lot with BoS and spikeBos, maybe helps a little with hardPD, does not help with easyPD.

We can go even more granular and look at whether words on a trial is associated with more bonus * on that trial *.

```{r}
success_round <- rounds %>% filter(chat_cond=="chat") %>% select(playerId, gameId, gametype, payoff, repNum) %>% inner_join(chat_by_pair) %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(gameId, gametype, repNum) %>% summarize(mean_payout=mean(payoff), mean_chat=mean(is.chat), mean_words=mean(words))

ggplot(success_round, aes(x=as.factor(mean_chat), y=mean_payout, color=gametype))+geom_jitter(alpha=.1, width=.2, height=0)+stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=0), size=.5)+labs(x="Was there chat this round", y="Mean payout this round")

ggplot(success_round, aes(x=mean_words, y=mean_payout, color=gametype))+geom_jitter(alpha=.1)+geom_smooth(method="lm")+labs(x="Words said this round", y="mean payout this round")
```

There are some real outlier rounds where people say a lot, but mostly they say not that much even when they do say things. 
Talking is better than not talking for most trial types, but there's not much of a dose-response pattern.

# Chat and choices

Rather than collapsing over outcome (which has some randomness), we can look at per-round, which quadrant in the game people chose.

A number of people in the "chat" condition didn't use chat, so we sometimes 3 way split for used chat / didn't use chat / couldn't chat (in contrast to just spliting by intent to treat).

```{r}
outcome <- rounds %>% select(-playerId, -time,-payoff, -box_A, -box_B, -AA, -AB, -BA, -BB, -mix) %>%  
  pivot_wider(names_from=role, values_from=response) %>% 
  mutate(outcome=str_c(p1,p2))  %>% left_join(chat_by_pair) #%>% mutate(is.chat=ifelse(words>0, 1,0))

outcome_summary <- outcome %>% group_by(gametype, chat_cond, repNum,outcome) %>% tally() %>% group_by(gametype, chat_cond, repNum) %>% summarize(pct=n/sum(n), outcome=outcome)

```

## normal BoS

In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are bad for both.

```{r}
# outcome %>% filter(gametype=="BoS") %>%
#   mutate(outcome_parity=ifelse(outcome %in% c("AA", "BB"),1, 0)) %>% 
#   ggplot(aes(x=repNum, y=outcome_parity))+geom_jitter(width=0, height=.05)+facet_grid(.~chat_cond)+geom_smooth(method="lm")+
#   labs(title="BoS:fraction choosing AA or BB (good) outcomes", y="Good outcome?")

outcome %>% filter(gametype=="BoS") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AA", "BB"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.5))+
  labs(title="BoS:fraction choosing good outcomes",y="Good outcome?")
```

Near chance if you don't talk, far above chance if you do.

## spike BoS

In spike BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are bad for both. (same distribution for spikeBoS, just more skewed).

```{r}
outcome %>% filter(gametype=="spikeBoS") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AA", "BB"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.5))+
  labs(title="spike BoS:fraction choosing good outcomes", y="Good outcome?")
```

Actually talking is conducive to good outcomes.

## easyPD

In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.

```{r}
outcome %>% filter(gametype=="easyPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AA"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.25))+
  labs(title="easyPD:fraction choosing good outcomes", y="Good outcome?")

outcome %>% filter(gametype=="easyPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AB","BA"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.5))+
  labs(title="easyPD:fraction choosing AB or BA outcomes")

outcome %>% filter(gametype=="easyPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("BB"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.25))+
  labs(title="easyPD:fraction choosing BB outcomes")
```
Everyone is likely to choose the good outcome, but those who talk are more reliable at it (especially early on). No real consistency in what is chosen otherwise.

## hardPD

In hardPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. BA and AB are welfare maximizing.

```{r}
outcome %>% filter(gametype=="hardPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AA"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.25))+
  labs(title="hardPD:fraction choosing AA outcomes")

outcome %>% filter(gametype=="hardPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AB","BA"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.5))+
  labs(title="hardPD:fraction choosing AB or BA outcomes")

outcome %>% filter(gametype=="hardPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("BB"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.25))+
  labs(title="hardPD:fraction choosing BB outcomes")

```
When chat isn't used, AA increases over the course of the game. When chat is used, AB/BA increases. These are mostly trading off with each other, the obviously bad BB option isn't selected much ever.

# When do people talk?

```{r}

game_chat <- chat %>% filter(chat_cond=="chat") %>% filter(!is.na(targets)) %>% select(-type) %>% left_join(rounds) %>% 
  mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, gametype, repNum) %>% summarize(words=sum(words))

ggplot(game_chat, aes(x=repNum,y=words, color=gametype))+geom_jitter()

ggplot(game_chat, aes(x=reorder(gameId, words), y=words, color=gametype))+stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.4), size=.5)+theme(axis.text.x=element_blank())

game_chat %>% mutate(is.chat=ifelse(words>0,1,0)) %>% group_by(gameId, gametype) %>% summarize(pct_chat=mean(is.chat)) %>% ggplot(aes(x=reorder(gameId, pct_chat), y=pct_chat, color=gametype))+geom_jitter()+theme(axis.text.x=element_blank())
```
Seems to be substantial group to group variability in if they talk never/ sometimes/ always. Isn't fully driven by trial type.

How do we increase the rates of talkers!!??

# Does doing well in one type mean doing well in another??

For each game, look at fraction of each trial type where there did which quadrant and correlate??

```{r}
outcome <- rounds %>% select(-playerId, -time,-payoff, -box_A, -box_B, -AA, -AB, -BA, -BB, -mix) %>%  
  pivot_wider(names_from=role, values_from=response) %>% 
  mutate(outcome=str_c(p1,p2))  %>% left_join(chat_by_pair) #%>% mutate(is.chat=ifelse(words>0, 1,0))

outcome_summary <- outcome %>% group_by(gametype, chat_cond, gameId, outcome) %>% tally() %>% group_by(gametype, chat_cond, gameId) %>% summarize(pct=n/sum(n), outcome=outcome) %>% pivot_wider(names_from=c(gametype, outcome), values_from=pct, values_fill=0) %>% 
  mutate(good_BoS=BoS_AA+BoS_BB,
         good_spikeBoS=spikeBoS_AA+spikeBoS_BB,
         uneven_easyPD=easyPD_AB+easyPD_BA,
         uneven_hardPD=hardPD_AB+hardPD_BA)

library(GGally)

outcome_summary %>% filter(chat_cond=="chat")%>% ungroup() %>%  select(good_BoS, good_spikeBoS, uneven_easyPD, uneven_hardPD) %>% ggpairs()

outcome_summary %>% filter(chat_cond=="nochat")%>% ungroup() %>%  select(good_BoS, good_spikeBoS, uneven_easyPD, uneven_hardPD) %>% ggpairs()

```

# What do they say strategy is?

```{r}

exit %>% select(chat_cond, strategy) %>% arrange(chat_cond) %>% kable()
```