### 2024-05-20 code/analysis_may2024.Rmd
data and prelim analysis

---
title: "Mix of PD and BoS"
output:
  html_document: 
    toc: true
---

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"), fig.width=7, fig.height=3)
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
    fromJSON(flatten = T)
}

## Data import constants
data_location <- "data/new_study1_240520"

date_start <- lubridate::ymd("2024-05-20")

image_location <- "write-ups/images"

model_location <- "code/models"
```

```{r bonus}
## This is for determining participant bonuses using the version of data with PID

# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>%
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(bonus*4/3,2)) %>%
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
#   filter(bonus>0) %>% write_csv(here(data_location,"for_prolific.csv"))
# #
# cost <- read_csv(here(data_location, "player_payments.csv"))
# #
# sum(cost$cost)
```

```{r, include=F, eval=F}
d.treatments <- read_csv(here(data_location, "treatments.csv")) %>%
  rename(treatmentId = `_id`) %>%
  select(treatmentId, name) |>
  mutate(
    chat_cond = ifelse(str_detect(name, "no_chat"), "no_chat", "chat"),
    game_cond = "may2024"
  )

d.games <- read_csv(here(data_location, "games.csv")) %>%
  rename(gameId = `_id`) %>%
  left_join(d.treatments) %>%
  filter(createdAt >= date_start) %>%
  select(gameId, playerIds, game_cond, chat_cond)

d.gameplayers <- d.games %>%
  separate(playerIds, c("p1", "p2")) %>%
  pivot_longer(p1:p2, values_to = "playerId") %>%
  select(-name)

d.players <- read_csv(here(data_location, "players.csv")) %>%
  select(playerId = `_id`, role = data.role, data.bonus) %>%
  left_join(d.gameplayers)
```

```{r, include=F, eval=F}
d.chat.raw <- read_csv(here(data_location, "rounds.csv"), guess_max = 10000) %>%
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), "{}", data.chat)) %>%
  rename(row_id = `_id`) %>%
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>%
  select(-ends_with("response"), -ends_with("time"), -ends_with("payoff"), -ends_with("role")) %>%
  rename(data.gametype = data.type) %>% # VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  left_join(d.games) %>%
  write_csv(here(data_location, "raw_chat.csv"))

d.round_results.raw <- read_csv(here(data_location, "rounds.csv"), guess_max = 10000) %>%
  filter(createdAt >= date_start) %>%
  rename(data.gametype = data.type) %>% # VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  rename_with(~ gsub("room", "player", .x, fixed = T)) %>%
  rename_with(~ gsub("player", "player_", .x, fixed = T)) %>%
  rename_with(~ gsub("response", "_response", .x, fixed = T)) %>%
  rename_with(~ gsub("time", "_time", .x, fixed = T)) %>%
  select(-chat) %>%
  gather(key, value, starts_with("player")) %>%
  separate(key, into = c("blah", "playerId", "info")) %>%
  spread(info, value) %>%
  select(-blah) %>%
  select(-stageIds, -index, -createdAt, -submitted) %>%
  filter(!is.na(response)) %>%
  mutate(time = as.numeric(time) / 1000) %>%
  mutate(targets = map(targets, .f = ParseJSONColumn)) %>%
  unnest(targets) %>%
  mutate(
    image = gsub("/experiment/", "", image, fixed = TRUE),
    image = gsub(".jpeg", "", image, fixed = TRUE)
  ) %>%
  pivot_wider(names_from = label, names_prefix = "box_", values_from = image) %>%
  mutate(payoffs = map(payoffs, .f = ParseJSONColumn)) %>%
  unnest(payoffs) %>%
  mutate(
    AA = ifelse(role == "p1", AA.p1, AA.p2),
    AB = ifelse(role == "p1", AB.p1, AB.p2),
    BA = ifelse(role == "p1", BA.p1, BA.p2),
    BB = ifelse(role == "p1", BB.p1, BB.p2),
    type = ifelse(AB == 0 & BA == 0, "BoS", "PD")
  ) %>%
  left_join(d.games)

d.exit.survey <- read_csv(here(data_location, "player-inputs.csv")) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c("gameId")) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE))

d.complete <- d.round_results.raw %>%
  group_by(gameId, playerId) %>%
  summarize(max_rep = max(repNum)) %>%
  filter(max_rep == 39) %>%
  select(gameId, playerId) %>%
  write_csv(here(data_location, "complete.csv"))

d.chat <- d.chat.raw %>%
  inner_join(d.complete) %>%
  write_csv(here(data_location, "raw_chat.csv"))
d.round_results <- d.round_results.raw %>%
  inner_join(d.complete) %>%
  write_csv(here(data_location, "raw_results.csv"))
d.exit <- d.exit.survey %>%
  inner_join(d.complete) %>%
  write_csv(here(data_location, "exit.csv"))
```

# Are they human?

```{r bot}
exit <- read_csv(here(data_location, "exit.csv"))

exit %>%
  group_by(game_cond, chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes))
```

Yay, they mostly think they're playing with a human!!

Looks like we have 32 games in each condition (from less attrition than I anticipated).

```{r}
chat <- read_csv(here(data_location, "raw_chat.csv")) %>%
  select(gameId, text, type, repNum, playerId, chat_cond, gametype, targets)

rounds <- read_csv(here(data_location, "raw_results.csv")) %>%
  select(gameId, repNum, playerId, payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype, role)
```

# Before chat

```{r}
games <- rounds |>
  select(gameId, chat_cond) |>
  unique()

before_chat <- chat |>
  filter(is.na(targets)) |>
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, chat_cond) |>
  summarize(words = sum(words)) |>
  full_join(games)

ggplot(before_chat, aes(x = chat_cond, y = words)) +
  geom_jitter(width = .1, color = "blue", alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")
```

People do talk to each other a little in the pre-game chat time.

# Points earned

Note that BoS has a lower points range than PD because of what range of random numbers is selected.

```{r}
bonuses <- rounds %>%
  group_by(playerId, gameId, chat_cond, gametype) %>%
  summarize(payoff = mean(payoff))

ggplot(bonuses, aes(x = chat_cond, y = payoff, color = str_c(chat_cond))) +
  geom_jitter(width = .1, alpha = .1) +
  facet_grid(. ~ gametype) +
  scale_color_viridis(discrete = T) +
  labs(y = "mean_reward", x = "condition") +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 20)) +
  labs(y = "Mean reward / trial") +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14)
  )
```
Not seeing much of a difference in either category...

# During chat

```{r}
game_chat <- chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, gametype, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    mean_payout = mean(payoff),
    is_chat = ifelse(words > 0, 1, 0)
  )

when_chat <- game_chat %>%
  filter(chat_cond == "chat") |>
  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
  group_by(repNum, gametype) %>%
  summarize(pct_chat = mean(is.chat))

ggplot(when_chat, aes(x = repNum, y = pct_chat, color = gametype)) +
  geom_point() +
  geom_smooth(method = "lm") +
  coord_cartesian(ylim=c(0,1))+
  labs(title = "pct of games who said anything each round")

game_chat |> filter(chat_cond == "chat") |>
  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
  group_by(gameId, gametype) %>%
  summarize(pct_chat = mean(is.chat)) |> 
  ggplot(aes(x=reorder(gameId, pct_chat, mean), y=pct_chat, color=gametype))+geom_point()+
  theme(axis.text.x=element_blank())+
    coord_cartesian(ylim=c(0,1))
```

looks like people are more likely to talk in BoS than PD games. But there's not a lot of talking at all. Driven by group-differences -- some pairs never or almost never chatted even if they could -- others used chat a substantial fraction of the time.

## Per trial does more talking help

 - chat_0 is had access to chat and didn't use on that trial
 - chat_1 is used chat on that trial
 - nochat_0 did not have access to chat

```{r}
ggplot(game_chat, aes(x = str_c(chat_cond, "_", is_chat), y = mean_payout)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_grid(. ~ gametype) +
  geom_smooth(method = "lm")
```

Looks like using language is correlated with better outcome with BoS and slightly better outcome for PD. But if this language helping or "people who avail themselves of the option to use language are more competent".

Is there a dose-response relationship, or is one word enough?

```{r}
ggplot(game_chat |> filter(words > 0), aes(x = log(words), y = mean_payout)) +
  geom_point() +
  facet_grid(. ~ gametype) +
  geom_smooth(method = "lm")
```

Especially where we have more data, looks like one word is enough. Indicative of coordination rather than negotiation, probably?

## Per game does talking help?

One idea is that talking on some trials may set up coordination strategies that can then effectively be used on later trials without talking on those trials.

So we want to look at overall volume of talking (in words or in # of trial talked) as a predictor for performance, controlling for talk on that round?

```{r}
group_talk <- game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words), pct_chat = mean(is_chat))

overall_talk <- game_chat |> left_join(group_talk)
ggplot(overall_talk |> filter(chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  geom_jitter(alpha = .1) +
  geom_smooth(method = "lm") +
  facet_grid(. ~ gametype) +
  theme(legend.position = "bottom")
```

Looks like talking on *other* trials might help in BoS if you didn't talk on this specific trial? Maaaybe? But there's a real data-sparsity issue.

# Option selected

 - chat_0 is had access to chat and didn't use on that trial
 - chat_1 is used chat on that trial
 - nochat_0 did not have access to chat

```{r}
outcome <- rounds %>%
  select(gameId, repNum, response, chat_cond, gametype, role) |>
  pivot_wider(names_from = role, values_from = response) %>%
  mutate(outcome = str_c(p1, p2)) %>%
  left_join(game_chat)
```

## BoS

In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are 0 for both.

```{r}
outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")
```

Near chance if you can't talk, above chance if you could but don't,  far above chance if you do.

```{r}
outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = repNum, y = outcome_parity)) +
  facet_grid(~ str_c(chat_cond, "_", is_chat)) +
  stat_summary(fun.data="mean_cl_boot", geom="point")+
  geom_smooth(method = "lm", se = F) +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes", y = "Good outcome?")
```

## PD

In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.

```{r}
outcome %>%
  filter(gametype == "easyPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) 
```

Can get a reasonable option no matter what. 
```{r}
outcome %>%
  filter(gametype == "easyPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = repNum, y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot", geom="point", alpha=.5) +
  facet_grid(.~str_c(chat_cond, "_", is_chat))+
  geom_smooth(method="lm")+
  geom_hline(aes(yintercept = .5)) 
```

# Language

How much language?

Filter only for games that talked at least a little.

Second graph filters for *trials* that talked.

```{r}
game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words)) |>
  filter(total_words != 0) |>
  left_join(game_chat) |>
  ggplot(aes(x = repNum, y = words)) +
  geom_point(alpha = .1) +
  geom_smooth()

game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words)) |>
  filter(total_words != 0) |>
  left_join(game_chat) |>
  filter(words > 0) |>
  ggplot(aes(x = repNum, y = words)) +
  geom_point(alpha = .1) +
  geom_smooth()
```

## how often do *both* people talk

```{r}
pair_chat <- chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, playerId, gametype, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    is_chat = ifelse(words > 0, 1, 0)
  ) |>
  group_by(gameId, gametype, chat_cond, repNum) |>
  summarize(how_many_chat = sum(is_chat)) |>
  group_by(gameId, gametype, chat_cond) |>
  summarize(pct_both = mean(how_many_chat == 2)) |>
  filter(chat_cond == "chat")

ggplot(pair_chat, aes(x = gametype, y = pct_both)) +
  geom_jitter(width = .2) +
  stat_summary(fun.data = "mean_cl_boot", color = "red")
```
Even in games that talk, there aren't *that* many trials where both people talk?

# Models

```{r}
game_chat <- chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, gametype, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    mean_payout = mean(payoff),
    is_chat = ifelse(words > 0, 1, 0)
  )

never_chat <- game_chat |>
  group_by(gameId, chat_cond) |>
  summarize(words = sum(words)) |>
  mutate(use_chat = ifelse(words > 0, 1, 0))

for_mod <- rounds |>
  select(gameId, repNum, playerId, payoff, chat_cond, gametype) |>
  left_join(game_chat) |>
  left_join(never_chat) |>
  mutate(
    has_chat = ifelse(chat_cond == "chat", 1, 0),
    is_BoS = ifelse(gametype == "BoS", 1, 0)
  ) # recode with no-chat PD as baseline
```

## Condition assignment on performance

```{r}
perf_priors <- c(
  set_prior("normal(5,2)", class = "Intercept"),
  set_prior("normal(0, 2)", class = "b"),
  set_prior("normal(0, 2)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

perf_mod <- brm(
  payoff ~ has_chat * is_BoS + repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_mod,
  prior = perf_priors,
  file = here(model_location, "perf_expt1"),
  control = list(adapt_delta = .95)
)

summary(perf_mod)
```

Note the high group level variability...

## Words over time

log(words) ~ game_type * trial_number

```{r}
for_log_words <- for_mod |>
  filter(words > 0) |>
  mutate(log_words = log(words))

lang_priors <- c(
  set_prior("normal(5,5)", class = "Intercept"),
  set_prior("normal(0, 5)", class = "b"),
  set_prior("normal(0, 5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

log_words_mod <- brm(
  log_words ~ is_BoS * repNum +
    (is_BoS * repNum | gameId),
  data = for_log_words,
  file = here(model_location, "words_expt1"),
  prior = lang_priors,
  control = list(adapt_delta = .95)
)
summary(log_words_mod)
```

More talk in BoS

## Words on outcome

outcome ~ game_type * log(words)

```{r}
perf_word_mod <- brm(
  payoff ~ log_words * is_BoS + repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_log_words,
  prior = perf_priors,
  file = here(model_location, "perf_words_expt1"),
  control = list(adapt_delta = .99)
)

summary(perf_word_mod)
```

Again, big group level variation.

# TODOs

* Look at the actual language

* Look at what people said their strategies were